\documentclass[titlepage, hidelinks, 12pt]{article}


\usepackage{lipsum}
\usepackage{hyperref}
\usepackage{palatino}

\usepackage{chngcntr}
\counterwithin{figure}{section}

\usepackage{setspace}
%\usepackage{indentfirst} %tex default is no indent on first paragraph after heading
\usepackage{url}
\usepackage{amsmath, amssymb, amsfonts, amsthm}
\usepackage{float}
\usepackage{subfig}
\usepackage{graphicx}
\usepackage{environ, enumerate}
%\usepackage{mathbbol}
%\DeclareSymbolFontAlphabet{\amsmathbb}{AMSb}
\graphicspath{ {images/} }
\providecommand{\keywords}[1]{\textbf{\textit{Keywords---}} #1} 
\usepackage[format=plain,
            labelfont={bf, it},
            textfont=it]{caption}

%%%%%%%%%
% indentation
%%%%%%%%%

\setlength\parindent{0pt}
\setlength{\parskip}{\baselineskip}

\setlength{\voffset}{-1cm}
\setlength{\textwidth}{17cm}
\addtolength{\textheight}{2cm}
\setlength{\footskip}{1cm}
\addtolength{\oddsidemargin}{-2cm}
\addtolength{\evensidemargin}{-2cm}

\widowpenalty10000
\clubpenalty10000

%Def, Lemma, Theorem, Corollary environment
\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[subsection]
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
%\newtheorem*{proof}{Proof}
\theoremstyle{remark}
\newtheorem*{remark}{Remark}
\newtheorem*{example}{Example}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}

%New commands
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\R}{\mathbb{R}}

%New math operators
\DeclareMathOperator{\diag}{diag}

\begin{document}


\title{Some metrics on the space of cancer treatments}
\author{PIMS $\text{Math}^{\text{Industry}}$ Team 3 (I think)}
\maketitle






\newpage

\section{Preamble and some notation}
Let the universal drug set be $\mathcal{D} = \left\{ d_1, d_2, \ldots \right\}$. Some of these drugs may be prescribed to treat a cancer: 
a specific cancer treament is some  $T = \left\{ d_{i_1}, \ldots, d_{i_n} \right\}\subseteq \mathcal{D}$. 
The set of all treatments is $\mathcal{T} = \mathcal{P}(\mathcal{D})$, the power set of $\mathcal{D}$. 

In an ideal test setting, an experimental treatment $T$ is given to a large sample and we check: did the treatment improve outcomes over
standard treatments? In practice, however, experimental treatments may be given imperfectly. That is, a treatment $T' \approx T$ may be given,
where $\approx$ is a vague notion of similar treatment.

One purpose of this project is to define a metric $d$ on $\mathcal{T}$ quantifying similarity of treatment. Or perhaps it isn't a metric that
we want\dots it probably makes statistical analysis easier if $d(T, T) = 1$ for any $T$, and if $d(T, T')$ gets smaller as $T'$ differs more
from $T$. So the Jaccard index $J(A, B) = |A\cap B|/|A\cup B|$ would be a reasonable first guess. 

\section{Matching score}
A confounding factor is that many drugs can target the same aberration. That is, while aionics might recommend drug $d$, a physician may
prescribe $d'$ without issue.  The Jaccard index is then probably not the greatest scoring available. Instead, we'll want to collapse drugs
down to the thing (this is vague, but I don't know the correct term) they target. 

The Matching Score basically does this. Vaguely, the matching score is defined by:

\begin{equation}
    \text{Matching Score }:= \frac{\text{\# of aberrations addressed by prescribed drug regimen}}{\text{\# of aberrations in cancerous tissue}}
    \label{eqn:matchingScore}
\end{equation}

Of course, the exact number of aberrations present and addressed is sometimes ambiguous, and certain drugs work better together than others.
The paper ``Molecular profiling of cancer patients enables personalized combination therapy: the I-PREDICT study'' runs through some of the
considerations when assigning a matching score. Bo quotes a similar (though not identical) description on Slack. 













\section{Considerations when constructing a score}
Drug choice seems to be one layer of abstraction too far. Rather, when evaluating how closely a drug regimen comes to desired, what we
wish to understand is: how closely does the regimen attack the issues present in the cancerous tissue? Further, how intensely shoudld we attack?
If I remember correctly, aionics provides the following. To each cancer profile $C$ is associated a symptom/treatment table: 

\begin{table}[h]
    \centering
\begin{tabular}{l|l|l|l}
aberration           & intensity       & importance   & drug options            \\\hline
name\_of\_aberration & $x\in\R_{\ge0}$ & $w\in(0, 1)$ & $d_1, d_2, \ldots, d_k$ \\
\end{tabular}
\end{table}

We can construct a similar table (first two columns, at least) for each drug regimen $T$. 
The question, now, is what is a reasonable scoring function $\sigma(C, T)$? Here $\sigma(C, T)$ is a scoring of closeness, evaluating
how well treatment $T$ attacks cancer profile $C$. 

As a first step, consider how well a regimen attacks a specific aberration. If we want to attack an aberration with
intensity $x$, and drug combination $T$ attacks with intensity $x_T$, then we'll let $\nu(x, x_T)$ be the intensity adjustment (adjusting
treatment effectiveness for how close to prescribed intensity we attack). For instance, we could let
\begin{equation}
    \nu(x, x_T) := e^{-(x-x_T)^2}.
    \label{eqn:intensityAdjustment}
\end{equation}
From here, it makes sense to define the score:

\begin{equation}
    \sigma(C, T) = \sum\limits_{a\in\text{Aberrations}} w_a\cdot \nu(x_a, x_{a,T})
    \label{eqn:ourMatchingScore}
\end{equation}

\newpage
\section{Matching score using graphs}
We know the aionic therapy identifies a subnetwork of genes that is the most active for each patient. This in turn we can think as a subgraph and derive our analysis from graph theory. 

An important measure that comes into place is the \textit{first Betti number} or \textit{circuit rank}, defined as $|E|-|V|+|C|$, where $E$ is the set of edges, $V$ the set of vertices and $C$ the set of connected components. In a given graph, we define a connected component as a maximal subgraph that satisfies that any two vertices are connected by a path. 

In our problem at hand, by the aiomic algorithm we have identified a certain subgraph of the gene map, let's call it $G$. From there, the proposed therapy would try to minimize the circuit rank of the subgraph generated by subtracting the vertices and edges from $G$ that the proposed drugs target. For a given therapy $T$ we denote this number by $B_T$.

In this context, it seems natural to try to quantify the change of the circut rank of the subgraph generated by subtracting the vertices and edges that correspond to the targeted genes of the actual therapy $T'$ that the patient receives. We could propose as matching score the quantity $B_{T'}/B_T$. 

\textbf{Potential problem with this score:} We know that drugs in general target more than only one gene. Potentially, this gene could very well be outside of our principal subgraph $G$, but could have interactions. How can we incorporate this into our matching score? 





\end{document}



